quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-13 paper-quollr.Rmd
Visualize the triangular mesh

To visualize the $2\text{-}D$ model, mainly three functions are used. As shown in Figure \@ref(fig:mesh-plots)a, `geom_trimesh()` to visualize the triangular mesh by adding a new layer to `ggplot()`. After identifying benchmark value to remove long edge, `vis_lg_mesh()` is used to visualize the triangular mesh by coloring the small and long edges. As shown in Figure \@ref(fig:mesh-plots)b, the small and long edges are colored by black and red respectively. Following this, `vis_rmlg_mesh()` is used to visualize smoothed $2\text{-}D$ model which is the $2\text{-}D$ model after removing the long edges (see Figure \@ref(fig:mesh-plots)c). In `vis_lg_mesh()` and `vis_rmlg_mesh()`, `benchmark_value` argument controls the edge removal in $2\text{-}D$. Using small value of `benchmark_value`, will produce a triangular mesh with missing data structure; for example `benchmark_value = 0.3` shows two clusters rather than continuous structure, while `benchmark_value = 0.8` creates long edges and mislead the data structure in $p\text{-}D$ space.

@@

>>>>

The `geom_trimesh()` function introduces a custom `ggplot2` layer designed for visualizing $2\text{-}D$ wireframe on a provided set of bin centroids.

To display the complete wireframe, users should supply all available bin centroids.

@@

```{r, echo=TRUE}
ggplot() + 
  geom_trimesh(
    data = df_bin_centroids, 
    aes(x = c_x, y = c_y)
    ) 
```

### 2025-05-13 paper-quollr.tex
Visualize the triangular mesh

To visualize the \(2\text{-}D\) model, mainly three functions are used. As shown in Figure \ref{fig:mesh-plots}a, \texttt{geom\_trimesh()} to visualize the triangular mesh by adding a new layer to \texttt{ggplot()}. After identifying benchmark value to remove long edge, \texttt{vis\_lg\_mesh()} is used to visualize the triangular mesh by coloring the small and long edges. As shown in Figure \ref{fig:mesh-plots}b, the small and long edges are colored by black and red respectively. Following this, \texttt{vis\_rmlg\_mesh()} is used to visualize smoothed \(2\text{-}D\) model which is the \(2\text{-}D\) model after removing the long edges (see Figure \ref{fig:mesh-plots}c). In \texttt{vis\_lg\_mesh()} and \texttt{vis\_rmlg\_mesh()}, \texttt{benchmark\_value} argument controls the edge removal in \(2\text{-}D\). Using small value of \texttt{benchmark\_value}, will produce a triangular mesh with missing data structure; for example \texttt{benchmark\_value\ =\ 0.3} shows two clusters rather than continuous structure, while \texttt{benchmark\_value\ =\ 0.8} creates long edges and mislead the data structure in \(p\text{-}D\) space.

@@

@@

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-31-1}

>>>>

The \texttt{geom\_trimesh()} function introduces a custom \texttt{ggplot2} layer designed for visualizing \(2\text{-}D\) wireframe on a provided set of bin centroids.

To display the complete wireframe, users should supply all available bin centroids.

@@

\begin{verbatim}
ggplot() + 
  geom_trimesh(
    data = df_bin_centroids, 
    aes(x = c_x, y = c_y)
    ) 
\end{verbatim}

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-24-1}

@@

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-32-1}
